---
title: "STAT 5014 hw3"
author: "Owusu Noah"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
  html_document:
    df_print: paged
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsfonts}
- \usepackage{fancyhdr}
- \usepackage{booktabs}
- \pagestyle{fancy}
- \fancyhf{}
- \rhead{Owusu Noah}
- \lhead{STAT PACKAGES hw3}
- \cfoot{\thepage}
- \usepackage{algorithm}
- \usepackage[noend]{algpseudocode}
- \setlength{\headheight}{13.59999pt}
geometry: margin=1in
spacing: single
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "!H", out.extra = "", out.width = "80%")

library(knitr)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(readr)
library(tidyr)
```

**Part A**

The issue associated with the thickness guage data is that each operator column appears trice and it is a duplication issue that requires tidying of the data.

```{r}
#importing thickness guage data

dat <- read.csv(file = "Thickness.csv")
pivot_dat <- dat %>%
  pivot_longer(cols = c("opt1","opt2","opt3"), names_to = "Operator", 
               values_to = "Thickness") #making the data more tidy
```

```{r}
#creating table for few observations
kable(head(pivot_dat), align = "cr", 
      caption = "First six rows of the Thickness dataset")
```


```{r}
#obtaining summary statistics of the data
s1 <- pivot_dat %>%
  group_by(Part) %>%
  get_summary_stats(Thickness, show = c("mean", "median", "min", "max"))

kable(head(s1), align = "cr", 
      caption = "Summary statistics of the Thickness dataset across Operator")

s2 <- pivot_dat %>%
  group_by(Operator) %>%
  get_summary_stats(Thickness, show = c("mean", "median", "min", "max"))
kable(head(s2), align = "cr", 
      caption = "Summary statistics of the Thickness dataset across Part")
```

```{r}
# Creating Box Plot for the data
ggplot(pivot_dat, aes(x = factor(Part), y = Thickness, fill = factor(Operator))) +
  geom_boxplot() +
  labs(title = "Box Plot of Thickness Guage  by Part with Operator")
```


**PART B**

Similar to part (A), the Body and Brain weight data has duplication issues. There are also missing values in this data.

```{r}
# Importing the Body and Brain weight data
bbw <- read.csv("BodyBrain_wt.csv")

#changing colum names
colnames(bbw) = c("Body", "Brain", "Body", "Brain", "Body", "Brain")

bd <- as.vector(rbind(bbw$Body,bbw$Body.1,bbw$Body.2))
bn <- as.vector(rbind(bbw$Brain,bbw$Brain.1,bbw$Brain.2))

df <- data.frame("Body wt" = bd, "Brain wt" = bn)

p_df <- df %>%
  pivot_longer(cols = c("Body.wt","Brain.wt"), names_to = "variable", 
               values_to = "weight")
```

\newpage
```{r}
kable(head(p_df), align = "cr", 
      caption = "First six rows of the Body and Brain weight dataset")
```


```{r}
#obtaining summary statistics
df1 <- p_df %>%
  group_by(factor(variable)) %>%
  get_summary_stats(weight, show = c("mean", "median", "min", "max"))
kable(head(df1), align = "cr", 
      caption = "Summary statistics of the Body and Brain weight dataset")
```


```{r}
# Create scatter plot for body and brain weight data
new_df <- df %>% 
  mutate(across(c(Body.wt, Brain.wt), function(x) log(x)))

new_df <- na.omit(new_df)

ggplot(new_df, aes(x = Brain.wt, y = Body.wt))+
geom_point(alpha = 0.5, col = "blue")+
geom_smooth(method = "lm", se = F, formula = y~x, col = "red")+
ggtitle("Figure 1: Scatter plot of Brain wt vs. Body wt (with linear fit)")+
stat_cor(method = ('pearson'),col='blue')+
stat_regline_equation(label.x=4.5, label.y=335, output.type = "latex")+
labs(x = "Body wt", y = "Brain wt")+
theme_classic()
```



**PART C**

The Long Jump data have duplication issue. Each variable appears four times on the column. There are also missing values in this data.

```{r}
#importing long jump dataset
long <- read.csv("LongJump.csv")

l1 <- as.vector(rbind(long$Year,long$Year.1,long$Year.2))
l2 <- as.vector(rbind(long$Jump,long$Jump.1,long$Jump.2))

df <- data.frame("Year" = l1, "Jump" = l2)

ndf <- df %>%
  pivot_longer(cols = c("Year","Jump"), names_to = "variable", 
               values_to = "value")
kable(head(ndf), align = "cr", 
      caption = "First six rows of the Long Jump dataset")

```

```{r}
ndf1 <- ndf %>%
  group_by(factor(variable)) %>%
  get_summary_stats(value, show = c("mean", "median", "min", "max"))
kable(head(ndf1), align = "cr", caption = "Summary statistics of the LongJump dataset")
```


```{r}
#Creating a scatter plot for the long jump dataset
ggplot(df, aes(x = Year, y = Jump))+
geom_point(alpha = 0.5, col = "blue")+
geom_smooth(method = "lm", se = F, formula = y~x, col = "red")+
ggtitle("Figure 1: Scatter plot of Long Jump vs. Year (with linear fit)")+
stat_cor(method = ('pearson'),col='blue')+
stat_regline_equation(label.x=60, label.y=335, output.type = "latex")+
labs(x = "Year", y = "Jump")+
theme_classic()
```

\newpage
**PART D**

The issue associated with the tomato data is that the values are embedded between each row and each column.

```{r}
#importing the tomatoe dataset
tom <- read.csv(file = "tomato.csv")
kable(head(tom), align = "cr", 
      caption = "First six rows of the tomatoe dataset")
```
```{r}
t1 <- tom %>%
  group_by(Variety) %>%
  get_summary_stats(PlantDensity, show = c("mean", "median", "min", "max"))
kable(head(t1), align = "cr", caption = "Summary statistics of the LongJump 
      dataset across variety")
```


```{r}
t2 <- tom %>%
  group_by(Unit) %>%
  get_summary_stats(PlantDensity, show = c("mean", "median", "min", "max"))
kable(head(t2), align = "cr", 
      caption = "Summary statistics of the LongJump dataset across unit")
```

```{r}

#Creating Box Plot for tomatoe dataset
ggplot(tom, aes(x = factor(Variety), y = PlantDensity, fill = factor(Unit))) +
  geom_boxplot() +
  labs(title = "Box Plot of PlantDensity  by Variety with Unit")
```
\
**Part E**

In the Larvae counts data, the Age and the Treatment variables seem to be interacted and have embedded values.

```{r}
#importing Larve counts dataset
LV <- "LarvaeCounts.csv" %>%
read.csv()

lv_dat <- LV %>%
  arrange(desc(Counts))

kable(head(lv_dat), align = "cr", caption = "First six rows of Larvae Count 
      dataset arranged in a descending order of counts")
```

```{r}
dt <- lv_dat %>%
  group_by(Treatment,Age) %>%
  get_summary_stats(Counts, show = c("mean", "median", "min", "max"))
kable(head(dt), align = "cr", caption = "Summary of the Larvae count dataset 
      based on Age and Treatment")
```

```{r}
dt <- lv_dat %>%
  group_by(Block) %>%
  get_summary_stats(Counts, show = c("mean", "median", "min", "max"))
kable(head(dt), align = "cr", caption = "Summary of the Larvae count dataset 
      based on Block")
```


```{r}
#Creating barplot for Larvae counts dataset
ggplot(lv_dat, aes(x=factor(Treatment),y=Counts,fill=factor(Age))) +
geom_bar(position="dodge",stat='identity')+
ggtitle("Barplot of Larvae Counts vrs Treatment across Age")+
xlab("Treatment")+ ylab("Larvae Counts")+
scale_fill_manual(values = c("cyan2","orange"))
```

